\pkg{sdmTMB}: An \proglang{R} Package for Fast, Flexible, and User-Friendly Generalized Linear Mixed Effects Models with Spatial and Spatiotemporal Random Fields

### 2023-10-19 doc/paper-jss/sdmTMB-paper.Rnw
Edits

A fitted model summary can be viewed with \fct{print} or \fct{summary} and a set of basic ``sanity'' checks can be run with \fct{sanity}.
\fct{tidy} returns parameter estimates in standard data frame formats similar to \pkg{broom} \citep{broom}.
Others standard S3 methods are also available such as \fct{fixef}, \fct{confint}, and \fct{vcov}.

@@

As an alternative to the CRAN version, the development version can be installed with:

<<remotes, eval=FALSE, echo=TRUE>>=
# install.packages("remotes")
remotes::install_github("pbs-assess/sdmTMB")
@

@@

<<dog-binomial-mesh2, results='hide', message=FALSE, warning=FALSE, echo=TRUE, fig.cap="SPDE mesh (lines) combined with the trawl survey observations (points). The locations where lines intersect are referred to as ``vertices''. Finer meshes will be slower to fit but generally increase the accuracy of the SPDE approximation. A greater degree of control over the mesh construction can be achieved by using \\pkg{fmesher} or \\proglang{R}-\\pkg{INLA} directly and supplying the object to \\code{make\\_mesh()}.", fig.width=4.5, fig.asp=1, out.width="3in">>=
mesh_pcod2 <- make_mesh(
  pcod, c("X", "Y"),
  fmesher_func = fmesher::fm_mesh_2d_inla,
  cutoff = 8, # minimum triangle edge length
  max.edge = c(10, 40), # max triangle edge length; inner and outer meshes
  offset = c(10, 40) # inner and outer border widths
)
plot(mesh_pcod2)
@

@@

We can also fit a version that omits the spatial random field by setting \code{spatial = "off"}.
We will use the \fct{update} S3 method to refit the model while updating any specified arguments:

@@

We can test for spatial autocorrelation with a Moran's I test or with a visual inspection of the residuals calculated from the \fct{residuals} S3 method.
We omit example code for Moran's I calculation for brevity but include it in the code supplement.
% TODO: reword this?
We see strong evidence for spatial autocorrelation for the model without a random field (p $<$ 0.01) but no evidence for the model with a random field (p = 0.84) suggesting that residual spatial autocorrelation is alleviated by including the random field.

>>>>

A fitted model summary can be viewed with \fct{print} or \fct{summary} and a set of basic ``sanity'' checks can be run with \fct{sanity}.
\fct{tidy} returns parameter estimates in standard data frame formats similar to \pkg{broom} \citep{broom}.
Others standard methods are also available such as \fct{fixef}, \fct{confint}, and \fct{vcov}.

@@

Users who wish to automatically install suggested packages as well may wish to use

<<sdmTMB-install-suggest, eval=FALSE, echo=TRUE>>=
install.packages("sdmTMB", dependencies = TRUE)
@

As an alternative to the CRAN version, the development version can be installed with

<<remotes, eval=FALSE, echo=TRUE>>=
install.packages("remotes")
remotes::install_github("pbs-assess/sdmTMB", dependencies = TRUE)
@

@@

<<dog-binomial-mesh2, results='hide', message=FALSE, warning=FALSE, echo=TRUE, fig.cap="SPDE mesh (lines) combined with the trawl survey observations (points). The locations where lines intersect are referred to as ``vertices''. Finer meshes will be slower to fit but generally increase the accuracy of the SPDE approximation. A greater degree of control over the mesh construction can be achieved by using \\pkg{fmesher} or \\proglang{R}-\\pkg{INLA} directly and supplying the object to \\code{make\\_mesh()}.", fig.width=4.5, fig.asp=1, out.width="3in">>=
mesh_pcod2 <- make_mesh(
  pcod,
  xy_cols = c("X", "Y"),
  fmesher_func = fmesher::fm_mesh_2d_inla,
  cutoff = 8,
  max.edge = c(10, 40),
  offset = c(10, 40)
)
plot(mesh_pcod2)
@

@@

We can also fit a version that omits the spatial random field by setting \code{spatial = "off"}.
We will use the \fct{update} method to refit the model while updating any specified arguments:

@@

We can test for spatial autocorrelation with a Moran's I test or with a visual inspection of the residuals calculated from the \fct{residuals} method.
We omit example code for Moran's I calculation for brevity but include it in the code supplement.
% TODO: reword this?
We see strong evidence for spatial autocorrelation for the model without a random field (p $<$ 0.01) but no evidence for the model with a random field (p = 0.84) suggesting that residual spatial autocorrelation is alleviated by including the random field.
